---
title: "The allometry of field metabolic rates in tropical vs. temperate hummingbirds"
author: "Anusha Shankar, Donald R. Powers, Catherine H. Graham"
date: "March, 2018"
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
  word_document: default
---
## Abstract
Allometric scaling relationships describe how the scaling of one trait (e.g. body size) affects the scaling of another trait (e.g. field metabolic rate, brain size). These scaling relationships have intrigued scientists for decades because there seem to be rules that govern how metabolic processes scale up with body size.  In birds, field metabolic rate (FMR) scales with body mass (M) in the form FMR = aM^0.67^. This relationship between metabolic rate and mass has been found to vary depending on the taxonomic level being studied. Thus, a scaling exponent on the scale of all birds could mask localized taxon-specific patterns. Hummingbirds have among the highest mass-specific rates of all vertebrates, as well as a much higher wing aspect ratio than predicted for their size. We collected FMR and mass data for hummingbirds, and examined the scaling of FMR with body mass among hummingbirds. We found that hummingbirds have an FMR to mass scaling exponent >0.9 -- much higher than any group previously studied. Further, since temperate hummingbirds tend to feed on densely clumped resources more than tropical hummingbirds tend to, we expect tropical birds to have a higher FMR to body size exponent than temperate birds to support increased foraging effort. We find that tropical hummingbirds have a higher FMR to mass ratio on average than temperate birds.

Tropical birds have been found to have significantly lower BMR than temperate birds, though within the tropics (Peru) elevation had no effect on BMR (Londono et al. 2015). 'Pace of life' is thus slower in tropical than temperate regions.

##Introduction:

```{r, echo=F}
## Maybe compare SC pre-monsoon birds instead of what Joey's poster does with HC pre-monsoon birds. Because temperatures were similar for SC and Maqui/SL. HC was colder.
## Calculate nighttime energy expenditure for the birds and subtract that from total DEE, what is daytime energy expenditure? What is hourly daytime energy expenditure?
## Earlier methods used jugular blood samples. Used to kill the bird.
## There are 3 DLW hummingbird papers. One of them in the tropics- Weathers and Stiles. Don's done the other two. Harry Teebo, captive DLW. 
## Austin Spence from Morgan Tingley's lab - toe-clipping to get haemoglobin. Same guy who learned respirometry from Don.
## Could correct DLW data for thermoregulation. If you have TNZ data for the birds.
## To do: Include BBLH data in plots
## Have to talk about testing for dose sizes in AZ birds because they were fed and not injected.
## For 2013 DLW calculations, assumed the birds had total body water of 0.7 (i.e. 70% proportion of water to body mass). Speakman said this was okay.
## Read more about allometric relationships, exponents starting with BMR. Need to be able to explain why the slope is 1. Birds overall, FMR to mass slope is 0.68.
## Is there data comparing tropics vs. temperate for other bird species that I could analyze?
## To estimate BMR from RMR, reduce RMR by 25%. Ashoff and Pohl papers compare BMR and RMR
## Create allometric relationships for different foraging groups. Insectivores vs. granivores vs. nectarivores
## Doug - scope for activity and bmr- is there a correlation?
```

**Hypotheses**:  
1. Daily energy expenditure will increase if daily temperatures vary more widely  
 + Higher thermoregulatory costs (avg. daily temperatures are low) would contribute to increased DEE (both directly because of increased thermoregulatory costs, and indirectly because they would need to be more active to gain energy to deal with these increased costs)  
2. Tropical resident species have a higher DEE than temperate long distance migrants  
3. Territorial species 

```{r, echo=F}
library(ggplot2)
library(RColorBrewer)

#opts_knit$set(root.dir )
#setwd("/Users/anshankar/Dropbox/DLW_paper")
## Includes data from XXXX papers.
dlw <- read.csv("/Users/anshankar/Dropbox/DLW_paper/Data/DLW_data2.csv")
dlw <- dlw[dlw$Reasonable_not=="Y",]

my_theme <- theme_classic(base_size = 15) + 
  theme(panel.border = element_rect(colour = "black", fill=NA)) 
colourCount <- length(unique(dlw$Species))
getPalette <- colorRampPalette(brewer.pal(9, "Set3"))

## Aggregate data frame by mean.
dlw_mean <- data.frame()
dlw_mean <- aggregate(dlw$kJ_day, by=list(dlw$Species, dlw$Big_site), FUN="mean", na.omit=T)
dlw_something <- aggregate(dlw$Mass_g, by=list(dlw$Species, dlw$Big_site), 
                           FUN="mean", na.omit=T)
dlw_mean <- merge(dlw_mean, dlw_something, by = c("Group.1", "Group.2"))
names(dlw_mean) <- c("Species", "Region", "kJ_day", "Mass_g")

## Fuction to plot sample size
give.n <- function(x){
  return(c(y = mean(x), label = length(x)))
}

lm_eqn <- function(y, x){
  m <- lm(y ~ x);
  eq <- substitute(italic(y) == 
                     a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
                   list(a = format(coef(m)[1], digits = 2), 
                        b = format(coef(m)[2], digits = 2), 
                        r2 = format(summary(m)$r.squared, digits = 3)))
  as.character(as.expression(eq));                 
}
```

A log-log plot of hummingbird daily energy expenditure (kJ) vs. mass (g), including values from this study as well as from the literature. Colors represent species:
```{r, echo=F}
#Species-wise log-log plot of DEE vs. mass
dlw_sp_mean_log <- ggplot(dlw_mean, aes(log(Mass_g), log(kJ_day), label=Species)) + 
  geom_point(aes(col=Species), size=4) + stat_smooth(method = 'lm', alpha=0.1) + #geom_smooth(method=lm) + 
  my_theme + theme(legend.key.height = unit(1, "line")) + 
  scale_colour_manual(values = getPalette(colourCount)) +
  geom_text(hjust = 0, nudge_x = 0.03) +
  xlab("log(Mass)") + ylab("log(Energy expenditure (kJ/day))") + xlim(1,3.1) +
  geom_text(x = 1.5, y = 4.6, label = lm_eqn(log(dlw_mean$kJ_day), log(dlw_mean$Mass_g)), parse=T) +
  geom_text(x = 1.5, y = 5, col="red", label = lm_eqn(log(dlw_mean$kJ_day[dlw_mean$Species != "PAGI"]), 
                                             log(dlw_mean$Mass_g[dlw_mean$Species != "PAGI"])), parse=T)
plot(dlw_sp_mean_log, fig.align="center")
```

A plot of daily energy expenditure (kJ) vs. Mass (g), including literature and study values:
```{r, echo=F}
dlw_sp_mean <- ggplot(dlw_mean, aes(Mass_g, kJ_day, label=Species)) + 
  geom_point(aes(col=Species), size=4) + theme_bw() + geom_smooth(method=lm) +
  geom_text(x = 8, y = 120, label = lm_eqn(dlw_mean$kJ_day, dlw_mean$Mass_g), parse=T) +
  geom_text(hjust = 0, position=position_jitter(width=0.5)) +
  geom_text(x = 8, y = 140, col="red", label = lm_eqn(log(dlw_mean$kJ_day[dlw_mean$Species != "PAGI"]), 
                                             log(dlw_mean$Mass_g[dlw_mean$Species != "PAGI"])), parse=T) +
  scale_colour_manual(values = getPalette(colourCount)) +
  theme(legend.key.height = unit(1, "line")) + xlab("Initial mass (g)") + ylab("Daily Energy Expenditure (kJ/day)") + xlim(-1,19)
plot(dlw_sp_mean, fig.align="center")
```

A plot of daily energy expenditure (kJ) vs. Mass (g), with species' mean in circles and individuals measurements in triangles:
```{r, echo=F}
dlw_indiv_and_sp_mean <- ggplot(NULL, aes(log(Mass_g), log(kJ_day))) +
  geom_point(data=dlw, aes(Mass_g, kJ_day, fill=Species), size=2, shape=24, alpha=0.7) +
  geom_point(data=dlw_mean, aes(Mass_g, kJ_day, fill=Species), size=4, shape=21) + theme_bw() + 
  scale_colour_manual(values = getPalette(colourCount)) +
  scale_fill_manual(values = getPalette(colourCount)) +
  theme(legend.key.height = unit(1, "line")) #+ xlab("Initial mass (g)") + ylab("Daily Energy Expenditure (kJ/day)")
plot(dlw_indiv_and_sp_mean, fig.align="center", fig.width = 10)
```


Daily energy expenditure vs. Mass, colored by species:
```{r, echo=F}
dee_total <- ggplot(dlw, aes(Mass_g, kJ_day, group=Species, col=Species)) + 
  geom_smooth(method='lm',se = F, size=2) +
  geom_point(size=2, shape=16, alpha=0.7) + theme_bw() + 
  scale_colour_manual(values = getPalette(colourCount)) +
  ylab("Daily Energy Expenditure (kJ/day)") + xlab("Initial mass (g)")
plot(dee_total, fig.align="center")
```

Mass-corrected daily energy expenditure vs. Mass, colored by species:
```{r, echo=F}
dee_per_gram <- ggplot(dlw, aes(Mass_g, kJ_dayg)) + 
  geom_point(aes(col=Species), size=2, shape=16) + theme_bw() + #scale_color_brewer(palette="Paired") +
  scale_colour_manual(values = getPalette(colourCount)) +
  ylab("kJ per day*g") + xlab("Initial mass (g)")
plot(dee_per_gram, fig.align="center")
```

Daily energy expenditure collapsed by species vs. average mass of the species, and shapes denote region:
```{r, echo=F}
dee_indiv_log <- ggplot(dlw_mean, aes(log(Mass_g), log(kJ_day))) + 
  geom_point(aes(col=Region), size=5) + theme_bw() + xlab("Log(Mass (g))") +
  scale_shape_manual("Region\n", values=c(16, 3, 1, 2), labels=c("Arizona", "Chile", "Costa Rica", "Ecuador")) +
  scale_colour_manual(values = getPalette(colourCount)) +
  ylab("Log(kJ per day)")  + geom_smooth(method=lm) + 
  theme(strip.background = element_blank(), 
        panel.border = element_rect(colour = "black", fill=NA), legend.key.height=unit(2,"line"))
plot(dee_indiv_log, fig.align="center")
```

Just Ecuadorian birds:
```{r, echo=F}
dee_indiv_EC <- ggplot(subset(dlw,Big_site %in% "EC"), aes(log(Mass_g), log(kJ_day))) +
  geom_point(aes(col=Species), size=5, alpha=0.8) + theme_bw() + xlab("Log(Mass (g))") + 
  scale_colour_manual(values = getPalette(colourCount)) +
  ylab("Log(kJ per day)")  + geom_smooth(method=lm) + 
  geom_text(x = 5, y = 75, label = lm_eqn(dlw$kJ_day, dlw$Mass_g),data = subset(dlw,Big_site %in% "EC"), parse=T) +
  theme(legend.key.height=unit(1,"line"))
plot(dee_indiv_EC, fig.align="center")
```

Just Arizona birds
```{r, echo=F}
dee_indiv_AZ <- ggplot(subset(dlw,Big_site %in% "AZ"), aes(log(Mass_g), log(kJ_day))) +
  geom_point(aes(col=Species), size=5, alpha=0.8) + theme_bw() + xlab("Log(Mass (g))") + 
  scale_colour_manual(values = getPalette(colourCount)) +
  ylab("Log(kJ per day)")  + geom_smooth(method=lm) + 
  theme(legend.key.height=unit(1,"line"))
plot(dee_indiv_AZ, fig.align="center")
```

Just Costa Rican birds:
```{r, echo=F}
dee_indiv_CR <- ggplot(subset(dlw,Big_site %in% "CR"), aes(log(Mass_g), log(kJ_day))) +
  geom_point(aes(col=Species), size=5, alpha=0.8) + theme_bw() + xlab("Log(Mass (g))") + 
  scale_colour_manual(values = getPalette(colourCount)) +
  ylab("Log(kJ per day)") + geom_smooth(method=lm) + 
  theme(legend.key.height=unit(1,"line")) 
plot(dee_indiv_CR, fig.align="center")
```


```{r, echo=F}
fitAll <- lm(log(kJ_day) ~ log(Mass_g),data = dlw)
summary(fitAll)

fitEC <- lm(log(kJ_day) ~ log(Mass_g),data = dlw[dlw$Big_site=="EC",])
summary(fitEC)

fit_spMean <- lm(log(kJ_day) ~ log(Mass_g),data = dlw_mean)
summary(fit_spMean)

fit_try <- lm(log(kJ_day) ~ log(Mass_g) + Region, data=dlw_mean)
summary(fit_try)
```