---
title: "Daily energy expenditure"
author: "Anusha Shankar"
date: "Thursday, March 17, 2016"
output:
  word_document: default
  pdf_document:
    fig_caption: yes
---
Abstract:
Big sentence  
Methods  
Results  
Implication  
250 words

##Introduction:
**Problem**: Organisms manage to survive by balancing their energy budgets. They usually do this with the help of fat reserves to buffer them if conditions get difficult. But some small organisms do not have much fat, and must deal with the variable energetic demands their environment places on them. I want to know what strategies hummingbirds use to manage their tight energy budgets. Is there a difference in how much energy birds spend in the tropics vs. temperate environments. e.g. birds migrate to temperate places because tropical environments don't have as many bursts of flowering events. Migration trade-offs: high mortality during trade-offs, but big gains at the other end. Do tropical hummingbirds have a higher EE than temperate, because resources are distributed in the tropics in such a way that birds spend more time foraging. 

```{r, echo=F}
## Maybe compare SC pre-monsoon birds instead of what Joey's posted does with HC pre-monsoon birds. Because temperatures were similar for SC and Maqui/SL. HC was colder.
## Calculate nighttime energy expenditure for the birds and subtract that from total DEE, what is daytime energy expenditure? What is hourly daytime energy expenditure?
## Earlier methods used jugular blood samples. Used to kill the bird.
## There are 3 DLW hummingbird papers. One of them in the tropics- Weathers and Stiles. Don's done the other two. Harry Teebo, captive DLW. 
## Austin Spence from Morgan Tingley's lab - toe-clipping to get haemoglobin. Same guy who learned respirometry from Don.
## Could correct DLW data for thermoregulation. If you have TNZ data for the birds.
## To do: Include BBLH data in plots
## Have to talk about testing for dose sizes in AZ birds because they were fed and not injected.
## For 2013 DLW calculations, assumed the birds had total body water of 0.7 (i.e. 70% proportion of water to body mass). Speakman said this was okay.
## Read more about allometric relationships, exponents starting with BMR. Need to be able to explain why the slope is 1. Birds overall, FMR to mass slope is 0.68.
## Is there data comparing tropics vs. temperate for other bird species that I could analyze?
## To estimate BMR from RMR, reduce RMR by 25%. Ashoff and Pohl papers compare BMR and RMR
## Create allometric relationships for different foraging groups. Insectivores vs. granivores vs. nectarivores
```

**Hurdle**: To answer this question in a useful way, we need daily energy expenditure measured in the wild, on free-living hummingbirds, not just on animals living in a cage. And we need this information in a variety of environmental conditions. Unfortunately, hummingbirds are too small to carry physiological sensors or GPS tags.

**Solution**: I used the doubly labeled water method to collect daily energy expenditure data from hummingbirds. Combining these data with daily activity budgets and estimates of energy spent on each activity, I will build an estimated daily energy budget for hummingbirds in North and South America.

**Benefit**: Hummingbirds are small endotherms with very high energetic demands, so they must manage their energy budgets carefully. They are always on the razor edge between life and death. Once we understand how they manage energy in changing environments, we can extend this energy budget model to other endotherms, such as those with large fat reserves. This will allow us to understand how endotherms will respond energetically to rapidly changing environmental conditions.

**Hypotheses**:
1. Whole animal daily energy expenditure will increase as mass increases
2. Daily energy expenditure will increase if daily temperatures vary more widely
 + Higher thermoregulatory costs (avg. daily temperatures are low) would contribute to increased DEE (both directly because of increased thermoregulatory costs, and indirectly because they would need to be more active to gain energy to deal with these increased costs)
3. Tropical resident species have a higher DEE than temperate long distance migrants
4. Territorial species 

```{r, echo=F, fig.cap="blah"}
library(ggplot2)

setwd("C:\\Users\\ANUSHA\\Dropbox\\DLW_paper")
## Includes data from XXXX papers.
dlw <- read.csv("DLW_data2.csv")
dlw <- dlw[dlw$Reasonable_not=="Y",]

## Aggregate data frame by mean.
dlw_mean <- data.frame()
dlw_mean <- aggregate(dlw$kJ_day, by=list(dlw$Species, dlw$Big_site), FUN="mean", na.omit=T)
dlw_something <- aggregate(dlw$Initial_mass, by=list(dlw$Species, dlw$Big_site), 
                           FUN="mean", na.omit=T)
dlw_mean <- merge(dlw_mean, dlw_something, by = c("Group.1", "Group.2"))
names(dlw_mean) <- c("Species", "Region", "kJ_day", "Mass_g")

## Fuction tO plot sample size
give.n <- function(x){
  return(c(y = mean(x), label = length(x)))
}

lm_eqn <- function(table, y, x){
  m <- lm(y ~ x, table);
  eq <- substitute(italic(y) == 
                     a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
                   list(a = format(coef(m)[1], digits = 2), 
                        b = format(coef(m)[2], digits = 2), 
                        r2 = format(summary(m)$r.squared, digits = 3)))
  as.character(as.expression(eq));                 
}
```

Plot of daily energy expenditure per gram bird, colored by species
```{r, echo=F, fig.cap=""}
#Species-wise log-log plot of DEE vs. mass
dlw_sp_mean <- ggplot(dlw_mean, aes(log(Mass_g), log(kJ_day))) + 
  geom_point(aes(col=Species), size=4) + theme_bw() + geom_smooth(method=lm)
dlw_sp_mean
```


```{r, echo=F}
dee_per_gram <- ggplot(dlw, aes(Initial_mass, kJ_dayg)) + 
  geom_point(aes(col=Species), size=2, shape=16) + theme_bw() + scale_color_brewer(palette="Paired") +
  ylab("kJ per day*g") + xlab("Initial mass (g)")
plot(dee_per_gram, fig.align='left')

dee_total <- ggplot(subset(dlw,Reasonable_not %in% "Y"), aes(Initial_mass, kJ_day)) + 
  geom_point(aes(col=Species), size=2, shape=16) + theme_bw() + 
  scale_color_brewer(palette="Paired") + ylab("kJ per day") + xlab("Initial mass (g)")
plot(dee_total, fig.align='left')

## For BBLH paper
#dee_BBLH <- ggplot(subset(dlw,Species %in% "BBLH"), aes(Initial_mass, kJ_day)) + 
 # geom_point(aes(shape=Pre_post_monsoon), size=7) + theme_classic(base_size = 30) + 
 # scale_shape_manual("Season\n", values=c(20,3), 
  #                   labels=c("Pre-monsoon", "Post-monsoon"), breaks=c("Pre", "Post")) + 
#  theme(strip.background = element_blank(),
 #       panel.border = element_rect(colour = "black", fill=NA),
  #      legend.key.height=unit(3,"line")) + 
  #ylab("Daily energy expenditure (kJ/day)") + xlab("Initial mass (g)")
#dee_BBLH

##To do: all birds combined, and then on Ecuador and on AZ birds separately
dee_indiv_log <- ggplot(dlw, aes(log(Initial_mass), log(kJ_day))) + 
  geom_point(aes(shape=Big_site), size=5) + theme_classic(base_size=30) + xlab("Log(Mass (g))") +
  scale_shape_manual("Region\n", values=c(16, 3, 1), labels=c("Arizona", "Costa Rica", "Ecuador")) +
  ylab("Log(kJ per day)")  + geom_smooth(method=lm) + 
  theme(strip.background = element_blank(), 
        panel.border = element_rect(colour = "black", fill=NA), legend.key.height=unit(2,"line"))
plot(dee_indiv_log)

dee_indiv_EC <- ggplot(subset(dlw,Big_site %in% "EC"), aes(log(Initial_mass), log(kJ_day))) +
  geom_point(aes(shape=Species), size=5) + theme_classic(base_size=30) + xlab("Log(Mass (g))") + 
  ylab("Log(kJ per day)")  + geom_smooth(method=lm) + 
  theme(strip.background = element_blank(), 
        panel.border = element_rect(colour = "black", fill=NA), legend.key.height=unit(2,"line"))
dee_indiv_EC

dee_indiv_AZ <- ggplot(subset(dlw,Big_site %in% "AZ"), aes(log(Initial_mass), log(kJ_day))) +
  geom_point(aes(shape=Species), size=5) + theme_classic(base_size=30) + xlab("Log(Mass (g))") + 
  ylab("Log(kJ per day)")  + geom_smooth(method=lm) + 
  theme(strip.background = element_blank(), 
        panel.border = element_rect(colour = "black", fill=NA), legend.key.height=unit(2,"line"))
dee_indiv_AZ

dee_indiv_CR <- ggplot(subset(dlw,Big_site %in% "CR"), aes(log(Initial_mass), log(kJ_day))) +
  geom_point(aes(shape=Species), size=5) + theme_classic(base_size=30) + xlab("Log(Mass (g))") + 
  ylab("Log(kJ per day)")  + geom_smooth(method=lm) + 
  theme(strip.background = element_blank(), 
        panel.border = element_rect(colour = "black", fill=NA), legend.key.height=unit(2,"line"))
dee_indiv_CR
```

Plot of total daily energy expenditure per bird, colored by species (not mass-corrected)
```{r, echo=F, fig.cap="blah"}
fitAll <- lm(log(kJ_day) ~ log(Initial_mass),data = dlw)
summary(fitAll)

fitEC <- lm(log(kJ_day) ~ log(Initial_mass),data = dlw[dlw$Big_site=="EC",])
summary(fitEC)

fit_spMean <- lm(log(kJ_day) ~ log(Mass_g),data = dlw_mean)
summary(fit_spMean)

fit_try <- lm(log(kJ_day) ~ log(Mass_g) + Region, data=dlw_mean)
summary(fit_try)
```