---
title: "Torpor paper graphs"
author: "Anusha Shankar, Rebecca Schroeder"
date: "Wednesday, December 09, 2015"
output: pdf_document
---
These are results or a cross-continent study of torpor in hummingbirds. Data were collected from a total of eight individuals across five sites, with three species in Arizona, USA, and five species in the Andes of Ecuador.

```{r echo=FALSE}
knitr::source('Torpor_paper_Nov2015.R')
```

**Nighttime energy expenditure plots**



**Average hourly energy expenditure plots**



**Minimum hourly energy expenditure plots**



**Comparing average and minimum energy expenditure**



**Species- or site-specific plots with regression lines**



**Statistical tests**



```{r}
summary(cars)
```
**Figure 1**: The thermoneutral zone (TNZ). Tb = Body temperature; LCT = Lower critical temperature;
UCT = Upper critical temperature; BMR = Basal metabolic rate

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
